Appendix

### 2023-07-06 docs/index.qmd
Fix appendix

abstract: |
    Language non-selectivity is prominent feature of the bilingual lexicon. An instance of such non-selectivity is embodied by cognateness, which recent studies have suggested to facilitate vocabulary acquisition in bilingual toddlers, who show larger vocabulary sizes when their languages share many cognates, and to be acquired at earlier ages than non-cognates. The specific mechanisms behind such facilitation are unclear. We present an account of early bilingual vocabulary growth in which cognateness interacts with lexical frequency and language exposure to facilitate the word acquisition. We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals toddlers. We used Bayesian Explanatory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler. We found cognateness facilitation effect only in low-frequency words. Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.

fig-dpi: 1000

---

```{r setup}
#| label: setup
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------

library(patchwork)
library(gt)
library(ggplot2)

targets::tar_config_set(store = here::here("_targets"),
                        script = here::here("_targets.R"))

targets::tar_load_globals()

targets::tar_load(c(bvq_data, childes, items, 
                    participants, responses))

targets::tar_load(c(model_fit, model_summary, model_epreds,
                    model_convergence, model_fit_prior, 
                    model_ppcs, posterior_doe_summary,
                    model_draws, model_summary,
                    posterior_doe_draws))

targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables))

# process objects
items <- bvq_data$pool  |>  
    mutate(item = str_remove(item, "cat_|spa_")) |> 
    select(item, language, semantic_category, class) |> 
    right_join(items, by = c("item", "language"))

rope_interval <- c(lower = -0.1, upper = 0.1)

# set ggplot theme and colour palette ------------------------------------------

my_theme <- theme_minimal() +
    theme(panel.grid = element_line(colour = "grey",
                                    linetype = "dotted"),
          axis.line = element_line(colour = "black"),
          text = element_text(size = 12, colour = "black"),
          axis.text = element_text(colour = "black"))

theme_set(my_theme)

bayesplot::bayesplot_theme_set(my_theme + theme(panel.grid = element_blank()))

clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

set.seed(888)

options(ggplot2.ordinal.fill = clrs,
        ggplot2.ordinal.colour = clrs,
        ggplot2.discrete.fill = clrs,
        ggplot2.discrete.colour = clrs,
        ggplot2.continuous.fill = scale_color_gradient,
        ggplot2.continuous.colour = scale_color_gradient,
        knitr.graphics.error = FALSE) 
```

::: {.callout-tip}
Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
:::

# Data

# Outputs

>>>>

fig-dpi: 1000
---

```{r setup}
#| label: setup
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------
targets::tar_config_set(store = here::here("_targets"),
                        script = here::here("_targets.R"))

yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
thanks <- yaml_metadata$thanks

```

::: callout-tip
Data and materials are available at [OSF](https://osf.io/hy984/), and
code is available at
[GitHub](https://github.com/gongcastro/trajectories).
:::

`r abstract`

`r thanks`

# Links

![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height=100px}

![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height=100px}

![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height=100px}

{{< include ../README.md >}}

### 2023-07-06 manuscript/_appendix/_syllable-frequency.qmd
Fix appendix

```{r tbl-items-table}
#| label: tbl-syll-items
#| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish
#| echo: false
#| warning: false
#| message: false
items |> 
    distinct(te, item, language, meaning, xsampa, n_syll) |> 
    left_join(distinct(pool_tmp, te, language, item, label),
              by = join_by(te, item, language)) |> 
    pivot_wider(names_from = language,
                values_from = c(item, label, xsampa, n_syll),
                id_cols = c(te, meaning),
                names_repair = janitor::make_clean_names,
                values_fn = first) |> 
    sample_n(15) |> 
    gt() |> 
    cols_hide(c(te, item_catalan, item_spanish)) |> 
    tab_spanner("Catalan", matches("catalan")) |> 
    tab_spanner("Spanish", matches("spanish")) |> 
    cols_label(meaning = "Translation",
               label_catalan = "Item",
               xsampa_catalan = "X-SAMPA",
               n_syll_catalan = "Syllables",
               label_spanish = "Item",
               xsampa_spanish = "X-SAMPA",
               n_syll_spanish = "Syllables") |> 
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) |> 
    tab_style(cell_text(style = "italic"),
              cells_column_labels())
```

Most Catalan and Spanish words had two syllables, with Spanish words having three and four syllables more often than Catalan words. Less than 1% of the words included in the analyses presented in the main body of the manuscripts had five syllables. No words had more than five syllables (see @fig-syll-number). We extracted lexical frequencies from the English corpora in the CHILDES database [@macwhinney2000childes; sanchez2019childesdb]. Using the Catalan and Spanish corpora was not possible due to the low number of children and tokens included in the corpora.

>>>>

```{r tbl-items-table}
#| label: tbl-syll-items
#| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish
#| echo: false
#| warning: false
#| message: false
tbl_data <- items |> 
    distinct(te, item, language, meaning, xsampa, n_syll) |> 
    left_join(distinct(pool_tmp, te, language, item, label),
              by = join_by(te, item, language)) |> 
    pivot_wider(names_from = language,
                values_from = c(item, label, xsampa, n_syll),
                id_cols = c(te, meaning),
                names_repair = janitor::make_clean_names,
                values_fn = first) |> 
    sample_n(15) |> 
    select(-c(te, item_catalan, item_spanish)) |> 
    relocate(meaning, matches("catalan"), matches("spanish"))

col_names <- c("Translation", "Item", "X-SAMPA", "Syllables",
               "Item", "X-SAMPA", "Syllables")

tbl_data |> 
    kable(format = "latex", 
          digits = 3,
          booktabs = TRUE,
          col.names = col_names, 
          align = "lllcllc") 
```

Most Catalan and Spanish words had two syllables, with Spanish words having three and four syllables more often than Catalan words. Less than 1% of the words included in the analyses presented in the main body of the manuscripts had five syllables. No words had more than five syllables (see @fig-syll-number). We extracted lexical frequencies from the English corpora in the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb]. Using the Catalan and Spanish corpora was not possible due to the low number of children and tokens included in the corpora.

### 2023-07-06 manuscript/appendix.qmd
Add comments by DAV
Fix appendix

---
title: "Appendix"
csl: "../assets/apa7.csl"
bibliography: "../assets/references.bib"
link-citations: true
editor: source

@@

format:
  pdf:
    documentclass: article
    papersize: letter
    fontsize: 12pt
    linestretch: 1.5
    header-includes:
    - \usepackage{tipa}
    geometry:
    - top=1in
    - bottom=1in
    - left=1in
    - right=1in
    colorlinks: true
    linkcolor: DodgerBlue
    urlcolor: DodgerBlue
    citecolor: DodgerBlue
    link-citations: true
    biblio-config: false
    block-headings: false
    indent: true
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------

library(patchwork)
library(gt)
library(ggplot2)

@@

targets::tar_load(c(bvq_data, childes, items, 
                    participants, responses))

targets::tar_load(c(model_fit, model_summary, model_epreds,
                    model_convergence, model_fit_prior, 
                    model_ppcs, posterior_doe_summary,
                    posterior_doe_draws))

@@

# process objects
items <- bvq_data$pool  |>  
    mutate(item = str_remove(item, "cat_|spa_")) |> 
    select(item, language, semantic_category, class) |> 
    right_join(items, by = c("item", "language"))

# set ggplot theme and colour palette ------------------------------------------

my_theme <- function() {
    theme_ggdist() +
        theme(panel.grid = element_line(colour = "grey",
                                        linetype = "dotted"),
              axis.line = element_line(colour = "black"),
              text = element_text(size = 12, colour = "black"),
              axis.text = element_text(colour = "black"))
}

theme_set(my_theme())

bayesplot::bayesplot_theme_set(my_theme + theme(panel.grid = element_blank()))

set.seed(888)

options(knitr.graphics.error = FALSE)
```

>>>>

---
csl: "../assets/apa7.csl"
bibliography: "../assets/references.bib"
link-citations: true
editor: source

@@

format:
  arxiv-pdf:
    keep-tex: true  
    linenumbers: true
    doublespacing: false
    runninghead: "Cognate beginnings to bilingual lexical acquistion (Appendix)"
    include-in-header: 
      text: |
        \usepackage{tipa}
        \usepackage{booktabs}
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------
library(knitr)
library(kableExtra)
library(ggplot2)

@@

targets::tar_load(c(bvq_data, childes, items, model_summary,
                    participants, responses))

targets::tar_load(c(posterior_doe_summary, posterior_doe_draws))

@@

my_theme <- theme_minimal() +
    theme(panel.grid = element_line(colour = "grey",
                                    linetype = "dotted"),
          axis.line = element_line(colour = "black"),
          text = element_text(size = 12, colour = "black"),
          axis.text = element_text(colour = "black"))

theme_set(my_theme)
```
